Bow shocks, bow waves, and dust waves. I. Strong coupling limit

### 2019-03-07 papers/Papers1234/bs-bw-dw-01.tex
Start writing about the shell structure figure
Write lots of discussion for Paper I

\newcommand\AddressCRyA{Instituto de Radioastronom\'{\i}a y Astrof\'{\i}sica,
  Universidad Nacional Aut\'onoma de M\'exico, Apartado Postal 3-72,
  58090 Morelia, Michoac\'an, M\'exico}
\author[Henney \& Arthur]{
  William J. Henney \& S. Jane Arthur\\
  \AddressCRyA
}

@@

\begin{document}
\label{firstpage}
\pagerange{\pageref{firstpage}--\pageref{lastpage}}
\maketitle
\begin{abstract}
  Dust waves and bow waves result from the action of a star's
  radiation pressure on a stream of dusty plasma that flows past it.
  They are an alternative mechanism to hydrodynamic bow shocks for
  explaining the curved arcs of infrared emission seen around some
  stars.  When gas and grains are perfectly coupled, for a broad class
  of stellar parameters, wind-supported bow shocks predominate when
  the ambient density is below \SIrange{100}{1000}{cm^{-3}}.  At
  higher densities radiation-supported bows can form, tending to be
  optically thin bow waves around B~stars, or optically thick bow
  shocks around early O~stars.  
\end{abstract}

>>>>

\newcommand\AddressCRyA{Instituto de Radioastronom\'{\i}a y Astrof\'{\i}sica,
  Universidad Nacional Aut\'onoma de M\'exico, Apartado Postal 3-72,
  58090 Morelia, Michoac\'an, M\'exico}
\author[Henney \& Arthur]{
  William J. Henney \& S. J. Arthur\\
  \AddressCRyA
}

@@

\begin{document}
\label{firstpage}
\pagerange{\pageref{firstpage}--\pageref{lastpage}}
\maketitle
\begin{abstract}
  Dust waves and bow waves result from the action of a star's
  radiation pressure on a stream of dusty plasma that flows past it.
  They are an alternative mechanism to hydrodynamic bow shocks for
  explaining the curved arcs of infrared emission seen around some
  stars.  When gas and grains are perfectly coupled, for a broad class
  of stellar parameters, wind-supported bow shocks predominate when
  the ambient density is below \SI{100}{cm^{-3}}.  At higher densities
  radiation-supported bows can form, tending to be optically thin bow
  waves around B~stars, or optically thick bow shocks around early
  O~stars.  For OB stars with particularly weak stellar winds,
  radiation-supported bows become more prevalent.
\end{abstract}

### 2019-03-07 papers/Papers1234/sec-strong-coupling.tex
Write lots of discussion for Paper I

\begin{figure*}
  \includegraphics[width=\linewidth]{figs/zones-v-n-plane}
  \caption{Bow regimes in parameter space (\(v, n\)) of the external
    stream for main-sequence OB stars of different masses:
    (a)~\SI{10}{M_\odot}, (b)~\SI{20}{M_\odot}, (c)~\SI{40}{M_\odot}.  In all
    cases, \(\kappa = \SI{600}{cm^2.g^{-1}}\) and efficient gas-grain
    coupling is assumed. Solid black lines of varying width show the
    bow size (star-apex separation, \(R_0\)), while gray shading shows
    the radiation bow wave regime, with lower border \(\tau = \eta\wind\) and
    upper border \(\tau = 1\), where \(\tau = 2 \kappa \rho R_0\) is the optical
    depth through the bow.  For bows above the red solid line, the
    ionization front is trapped inside the bow.  Blue lines delineate
    different cooling regimes.  Above the thin blue line
    (\(d_{\text{cool}} = h_0\)), the bow shock radiates efficiently,
    forming a thin shocked shell.  Below the thick blue line
    (\(d_{\text{cool}} = R_0\)), the bow shock is essentially
    non-radiative.}
  \label{fig:zones-v-n-plane}
\end{figure*}

@@

\begin{table*}
  \centering
  \caption{Stellar parameters for example stars}
  \label{tab:stars}
  \begin{tabular}{l S S S S S S S S S l}
    \toprule
    & {\(M / \si{M_\odot}\)} & {\(L_4\)}
    & {\(\dot{M}_{-7}\)} & {\(V_3\)} & {\( \eta\wind \)}
    & {Sp.~Type} 
    & {\(T_{\text{eff}} / \si{kK}\)} & {\(\lambda_{\text{eff}}\) / \si{\um}}
    & {\(S_{49}\)} & Figures 
    \\
    \midrule
    & 10 & 0.63 & 0.0034 & 2.47 & 0.0066 & {B1.5\,V} & 25.2 & 0.115 & 0.00013
                   & \ref{fig:zones-v-n-plane}a, \ref{fig:existence-dust-wave}a,
                     \ref{fig:phase-space-trajectories}--~\ref{fig:frozen-trajectories} \\
    Main-sequence OB stars
    & 20 & 5.45 & 0.492 & 2.66 & 0.1199 & {O9\,V} & 33.9 & 0.086 & 0.16
                   & \ref{fig:zones-v-n-plane}b, \ref{fig:existence-dust-wave}b \\
    & 40 & 22.2 & 5.1 & 3.31 & 0.4468 & {O5\,V} & 42.5 & 0.068 & 1.41
                   & \ref{fig:zones-v-n-plane}c, \ref{fig:existence-dust-wave}c \\[\smallskipamount]
    Blue supergiant star
    & 33 & 30.2 & 20.2 & 0.93 & 0.3079 & {B0.7\,Ia} & 23.5 & 0.123 & 0.016
                   & \ref{fig:B-supergiant} \\[\smallskipamount]
    Red supergiant star
    & 20 & 15.6 & 100 & 0.015 & 0.0476 & {M1\,Ia} & 3.6 & 0.805 & 0
                   & \ref{fig:M-supergiant} \\ 
    \bottomrule
  \end{tabular}
\end{table*}

We now consider the application to bow shocks around main sequence OB
stars, as well as cool and hot supergiants, expressing stellar and
ambient parameters in terms of typical values as follows:
\begin{align*}
  \label{eq:stellar-parameters}
  \dot{M}_{-7} &= \dot{M} / \bigl(\SI{e-7}{M_\odot.yr^{-1}}\bigr) \\
  V_3 &= V / \bigl(\SI{1000}{km.s^{-1}}\bigr) \\
  L_4 &= L / \bigl(\SI{e4}{L_\odot}\bigr) \\
  v_{10} &= v_\infty / \bigl( \SI{10}{km.s^{-1}} \bigr) \\
  n &= (\rho / \bar{m}) / \bigl( \SI{1}{cm^{-3}} \bigr) \\
  \kappa_{600} &= \kappa / \bigl( \SI{600}{cm^2.g^{-1}} \bigr) \ ,
\end{align*}
where \(\bar{m}\) is the mean mass per hydrogen nucleon
(\(\bar{m} \approx 1.3 m_{\text{p}} \approx \SI{2.17e-24}{g}\) for solar
abundances).  Note that \(\kappa = \SI{600}{cm^2.g^{-1}}\) corresponds to a
cross section of \(\approx \SI{e-21}{cm^2}\) per hydrogen nucleon, which is
typical for interstellar medium dust \citep{Bertoldi:1996a} at far
ultraviolet wavelengths, where OB stars emit most of their radiation.
In terms of these parameters, we can express the stellar wind momentum
efficiency as
  \label{eq:wind-eta-typical}
  \eta\wind = \num{0.495} \,\dot{M}_{-7} \,V_3  \,L_4^{-1}
and the fiducial radius and optical depth as
\begin{align}
  \label{eq:Rstar-typical}
  R_* / \si{pc} &= \num{2.21} \, (L_4 / n)^{1/2} \,v_{10}^{-1} \\
  \label{eq:taustar-typical}
  \tau_* &= \num{0.0089} \,\kappa_{600} \, (L_4 \,n)^{1/2} \,v_{10}^{-1} \ .
\end{align}
In Figure~\ref{fig:zones-v-n-plane}, we show results for the bow size
(apex distance, \(R_0\)) as a function of the density, \(n\), and
relative velocity, \(v_\infty\), of the external stream, with each panel
corresponding to a particular star, with parameters as shown in
Table~\ref{tab:stars}.  To facilitate comparison with previous work,
we choose stellar parameters similar to those used in the
hydrodynamical simulations of \citet{Meyer:2014b, Meyer:2016a,
  Meyer:2017a}, based on stellar evolution tracks for stars of
\SIlist{10;20;40}{M_\odot} \citep{Brott:2011a} and theoretical wind
prescriptions \citep{de-Jager:1988a, Vink:2000a}.  Although the
stellar parameters do evolve with time, they change relatively little
during the main-sequence lifetime of several million years.\footnote{%
  \label{fn:meyer-velocities-too-low}
  Note that we have recalculated the stellar wind terminal velocities,
  since the values given in the \citeauthor{Meyer:2014b} papers are
  troublingly low.  We have used the prescription
  \(V = 2.6 V_{\text{esc}}\), where
  \(V_{\text{esc}} = \left( 2 G M (1 - \Gamma_e)/ R \right)^{1/2}\) is the
  photospheric escape velocity, which is appropriate for strong
  line-driven winds with \(T_{\text{eff}} > \SI{21 000}{K}\)
  \citep{Lamers:1995a}.  We find velocities of
  \SIrange{2500}{3300}{km.s^{-1}}, which are consistent with
  observations and theory \citep{Vink:1999a} for O~stars, but at least
  two times higher than those cited by \citet{Meyer:2014b}. For
  main-sequence B~stars, wind column densities are too low to reliably
  measure the terminal velocity from near ultraviolet P~Cygni profiles
  \citep{Prinja:1989a}, and so the values are theory-dependent
  \citep{Krticka:2014a} and hence more uncertain.  A further
  complication is the existence of a subset of OB stars with
  anomalously weak winds \citep{Puls:2008a}, which in some cases is
  related to the presence of strong (\(\sim \SI{1}{kG}\)) magnetic fields
  \citep{Oskinova:2011b}.} %
The three examples are an early B~star (\SI{10}{M_\odot}), a late O~star
(\SI{20}{M_\odot}), and an early O~star (\SI{40}{M_\odot}), which cover the
range of luminosities and wind strengths expected from bow-producing
hot main sequence stars.  The luminosity is a steep function of
stellar mass (\(L \sim M^{2.5}\)) and the wind mass-loss rate is a steep
function of luminosity (\(\dot{M} \sim L^{2.2}\)), which means that the
wind momentum efficiency is also a steep function of mass
(\(\eta\wind \sim M^3\)), approaching unity for early O~stars, but falling to
less than 1\% for B~stars.

@@

In principle, gravitational attraction from the star, of mass \(M\),
will partially counteract the radiative acceleration.  This can be
accounted for by replacing \(L\) with an effective luminosity
\newcommand\Edd{\ensuremath{_{\text{E}}}}
\begin{equation}
  \label{eq:effective-luminosity}
  L_{\text{eff}} = L \bigl(1 - \Gamma\Edd^{\,-1}\bigr) \ ,
\end{equation}
in which \(\Gamma\Edd\) is the Eddington factor:
\begin{equation}
  \label{eq:eddington-factor}
  \Gamma\Edd = \frac{\kappa L}{4\pi c G M} = 458.5 \, \frac{\kappa_{600} L_4}{ M } \ ,
\end{equation}
where, in the last expression, \(M\) is measured in solar masses.  For
the stars in Table~\ref{tab:stars}, we find
\(\Gamma\Edd \approx \text{\numrange{30}{400}}\), so gravity can be safely
ignored.  The only exception is when the optical depth of the bow is
very large: \(\tau > \ln\Gamma\Edd \sim 5\), in which case gravity may be
important in the outer parts of the shell (see
\citealt{Rodriguez-Ramirez:2016b}).

@@

The bow shell consists of two parts: an inner zone of shocked stellar
wind, and an outer zone of decelerated ambient stream. For OB stars,
dust grains are present only in the ambient stream, whereas for cool
stars they will also be present in the stellar wind.  In the remainder
of this paper, we will concentrate on the OB star case, where it is
the outer zone that is most important observationally.  Bows are
typically detected via their infrared radiation (absorption and
re-emission by dust of stellar radiation), or by emission lines such
as the hydrogen H\(\alpha\) line.

@@

For the optically thick RBS case (Fig.~\ref{fig:axial-structure}c),
the shell density profile is no longer simply exponential since the
pressure scale height now increases as one moves away from the star,
due to the extinction-induced decline in radiation flux. Assuming
single scattering and plane geometry, the resultant hydrostatic
density profile is doubly exponential:
\begin{equation}
  \label{eq:rbs-density-profile}
  n(R) = n_2 e^{-e^{-x}} \text{\quad where\quad } x = n_2 \bar{m} \kappa (R - R_*) \ . 
\end{equation}
The shell density increases from \(e^{-1} n_2\) at \(R_*\) to saturate
at \(n_2\) for \(R > R_* + h_a\), where \(h_a \approx 8 h_0 / \tau\).  We show
results for a total shell optical depth \(\tau = 3\), for which
\(h_a\) and \(h_0\) are almost equal.  For larger optical depths,
\(h_0 > h_a\) so that the constant density portion of the shell will
dominate the total column.  A small fraction \(e^{-\tau}\) of the photons
will penetrate the shell and be available to decelerate the supersonic
stream, as in the RBW case, yielding
\(v_0 = (1 - \tau e^{-\tau}) v_\infty\).  In the illustrated case of
\(\tau = 3\), the resultant speed reduction before the shock is only 8\%.

@@

@@

@@

>>>>

\begin{figure*}
  \includegraphics[width=\linewidth]{figs/zones-v-n-plane}
  \caption{Bow regimes in parameter space (\(v, n\)) of the external
    stream for main-sequence OB stars of different masses:
    (a)~\SI{10}{M_\odot}, (b)~\SI{20}{M_\odot}, (c)~\SI{40}{M_\odot}.  In all
    cases, \(\kappa = \SI{600}{cm^2.g^{-1}}\) and efficient gas-grain
    coupling is assumed. Solid black lines of varying width show the
    bow size (star-apex separation, \(R_0\)), while gray shading shows
    the radiation bow wave regime, with lower border
    \(\tau = \eta\wind\) and upper border \(\tau = 1\), where
    \(\tau = 2 \kappa \rho R_0\) is the optical depth through the bow.  For bows
    above the red solid line, the ionization front is trapped inside
    the bow (\S~\ref{sec:trapp-ioniz-front}).  Blue lines delineate
    different cooling regimes (\S~\ref{sec:radi-cool-lengths}).  Above
    the thin blue line (\(d_{\text{cool}} = h_0\)), the bow shock
    radiates efficiently, forming a thin shocked shell.  Below the
    thick blue line (\(d_{\text{cool}} = R_0\)), the bow shock is
    essentially non-radiative.}
  \label{fig:zones-v-n-plane}
\end{figure*}

@@

\begin{table*}
  \centering
  \caption{Stellar parameters for example stars}
  \label{tab:stars}
  \begin{tabular}{l S S S S S S S S S l}
    \toprule
    & {\(M / \si{M_\odot}\)} & {\(L_4\)}
    & {\(\dot{M}_{-7}\)} & {\(V_3\)} & {\( \eta\wind \)}
    & {Sp.~Type} 
    & {\(T_{\text{eff}} / \si{kK}\)} & {\(\lambda_{\text{eff}}\) / \si{\um}}
    & {\(S_{49}\)} & Figures 
    \\
    \midrule
    & 10 & 0.63 & 0.0034 & 2.47 & 0.0066 & {B1.5\,V} & 25.2 & 0.115 & 0.00013
                   & \ref{fig:zones-v-n-plane}a \\
    Main-sequence OB stars
    & 20 & 5.45 & 0.492 & 2.66 & 0.1199 & {O9\,V} & 33.9 & 0.086 & 0.16
                   & \ref{fig:zones-v-n-plane}b, \ref{fig:O-weak-wind} \\
    & 40 & 22.2 & 5.1 & 3.31 & 0.4468 & {O5\,V} & 42.5 & 0.068 & 1.41
                   & \ref{fig:zones-v-n-plane}c\\[\smallskipamount]
    Blue supergiant star
    & 33 & 30.2 & 20.2 & 0.93 & 0.3079 & {B0.7\,Ia} & 23.5 & 0.123 & 0.016
                   & \ref{fig:B-supergiant} \\[\smallskipamount]
    Red supergiant star
    & 20 & 15.6 & 100 & 0.015 & 0.0476 & {M1\,Ia} & 3.6 & 0.805 & 0
                   & \ref{fig:M-supergiant} \\ 
    \bottomrule
  \end{tabular}
\end{table*}

We now consider the application to bow shocks around main sequence OB
stars, as well as cool and hot supergiants, expressing stellar and
ambient parameters in terms of typical values as follows:
\begin{align*}
  \label{eq:stellar-parameters}
  \dot{M}_{-7} &= \dot{M} / \bigl(\SI{e-7}{M_\odot.yr^{-1}}\bigr) \\
  V_3 &= V / \bigl(\SI{1000}{km.s^{-1}}\bigr) \\
  L_4 &= L / \bigl(\SI{e4}{L_\odot}\bigr) \\
  v_{10} &= v_\infty / \bigl( \SI{10}{km.s^{-1}} \bigr) \\
  n &= (\rho / \bar{m}) / \bigl( \SI{1}{cm^{-3}} \bigr) \\
  \kappa_{600} &= \kappa / \bigl( \SI{600}{cm^2.g^{-1}} \bigr) \ ,
\end{align*}
where \(\bar{m}\) is the mean mass per hydrogen nucleon
(\(\bar{m} \approx 1.3 m_{\text{p}} \approx \SI{2.17e-24}{g}\) for solar
abundances).  Note that \(\kappa = \SI{600}{cm^2.g^{-1}}\) corresponds to a
cross section of \(\approx \SI{e-21}{cm^2}\) per hydrogen nucleon, which is
typical for interstellar medium dust \citep{Bertoldi:1996a} at far
ultraviolet wavelengths, where OB stars emit most of their radiation.
In terms of these parameters, we can express the stellar wind momentum
efficiency as
\begin{equation}
  \label{eq:wind-eta-typical}
  \eta\wind = \num{0.495} \,\dot{M}_{-7} \,V_3  \,L_4^{-1}
\end{equation}
and the fiducial radius and optical depth as
\begin{align}
  \label{eq:Rstar-typical}
  R_* / \si{pc} &= \num{2.21} \, (L_4 / n)^{1/2} \,v_{10}^{-1} \\
  \label{eq:taustar-typical}
  \tau_* &= \num{0.0089} \,\kappa_{600} \, (L_4 \,n)^{1/2} \,v_{10}^{-1} \ .
\end{align}
In Figure~\ref{fig:zones-v-n-plane}, we show results for the bow size
(apex distance, \(R_0\)) as a function of the density, \(n\), and
relative velocity, \(v_\infty\), of the external stream, with each panel
corresponding to a particular star, with parameters as shown in
Table~\ref{tab:stars}.  To facilitate comparison with previous work,
we choose stellar parameters similar to those used in the
hydrodynamical simulations of \citet{Meyer:2014b, Meyer:2016a,
  Meyer:2017a}, based on stellar evolution tracks for stars of
\SIlist{10;20;40}{M_\odot} \citep{Brott:2011a} and theoretical wind
prescriptions \citep{de-Jager:1988a, Vink:2000a}.  Although the
stellar parameters do evolve with time, they change relatively little
during the main-sequence lifetime of several million years.\footnote{%
  \label{fn:meyer-velocities-too-low}
  Note that we have recalculated the stellar wind terminal velocities,
  since the values given in the \citeauthor{Meyer:2014b} papers are
  troublingly low.  We have used the prescription
  \(V = 2.6 V_{\text{esc}}\), where
  \(V_{\text{esc}} = \left( 2 G M (1 - \Gamma_e)/ R \right)^{1/2}\) is the
  photospheric escape velocity, which is appropriate for strong
  line-driven winds with \(T_{\text{eff}} > \SI{21 000}{K}\)
  \citep{Lamers:1995a}.  We find velocities of
  \SIrange{2500}{3300}{km.s^{-1}}, which are consistent with
  observations and theory \citep{Vink:1999a}, but at least two times
  higher than those cited by \citet{Meyer:2014b}. } %
The three examples are an early B~star (\SI{10}{M_\odot}), a late O~star
(\SI{20}{M_\odot}), and an early O~star (\SI{40}{M_\odot}), which cover the
range of luminosities and wind strengths expected from bow-producing
hot main sequence stars.  The luminosity is a steep function of
stellar mass (\(L \sim M^{2.5}\)) and the wind mass-loss rate is a steep
function of luminosity (\(\dot{M} \sim L^{2.2}\)), which means that the
wind momentum efficiency is also a steep function of mass
(\(\eta\wind \sim M^3\)), approaching unity for early O~stars, but falling
to less than 1\% for B~stars.

@@

In principle, gravitational attraction from the star, of mass \(M\),
will partially counteract the radiative acceleration.  This can be
accounted for by replacing \(L\) with an effective luminosity
\newcommand\Edd{\ensuremath{_{\text{E}}}}
\begin{equation}
  \label{eq:effective-luminosity}
  L_{\text{eff}} = L \bigl(1 - \Gamma\Edd^{\,-1}\bigr) \ ,
\end{equation}
in which \(\Gamma\Edd\) is the Eddington factor:
\begin{equation}
  \label{eq:eddington-factor}
  \Gamma\Edd = \frac{\kappa L}{4\pi c G M} = 458.5 \, \frac{\kappa_{600} L_4}{ M } \ ,
\end{equation}
where, in the last expression, \(M\) is measured in solar masses.  For
the stars in Table~\ref{tab:stars}, we find
\(\Gamma\Edd \approx \text{\numrange{30}{400}}\), so gravity can be safely
ignored.  When the optical depth of the bow is very large,
\(\tau > \ln\Gamma\Edd \sim 5\), gravity will exceed the radiation force in the
outer parts of the shell (see \citealt{Rodriguez-Ramirez:2016b}), but
it is generally too weak to effect the shell structure even in such a
case, see \S~\ref{sec:axial-structure-bow} below.

@@

There are two shocked zones in the bow: an inner zone of shocked
stellar wind, and an outer zone of decelerated ambient stream. For OB
stars, dust grains are present only in the ambient stream, whereas for
cool stars they will also be present in the stellar wind.  In the
remainder of this paper, we will concentrate on the OB star case,
where it is the outer zone that is most important observationally.
Bows are typically detected via their infrared radiation (absorption
and re-emission by dust of stellar radiation), or by emission lines
such as the hydrogen H\(\alpha\) line.

@@

For the optically thick RBS case (Fig.~\ref{fig:axial-structure}c),
the shell density profile is no longer simply exponential since the
pressure scale height now increases as one moves away from the star,
due to the extinction-induced decline in radiation flux. Assuming
single scattering and plane geometry, the resultant hydrostatic
density profile is doubly exponential:
\begin{equation}
  \label{eq:rbs-density-profile}
  n(R) = n_2 e^{-e^{-x}} \text{\quad where\quad } x = n_2 \bar{m} \kappa (R - R_*) \ . 
\end{equation}
The shell density increases from \(e^{-1} n_2\) at \(R_*\) to saturate
at \(n_2\) for \(R > R_* + h_a\), where \(h_a\) is now the pressure
scale height at the inner edge of the shell.  From
equations~(\ref{eq:Rstar}, \ref{eq:rad:R0}, \ref{eq:tau-thin},
\ref{eq:strong-cooling-h0}, \ref{eq:scale-height}), and taking \(R_0 \approx R_*\), we find
\begin{equation}
  \label{eq:ha-versus-h0}
  h_a \approx \frac{4 h_0}{3 \tau} \ .
\end{equation}
We show results for a total shell optical depth \(\tau = 3\), for which
\(h_a\) is a significant fraction of \(h_0\).  For much larger optical
depths, \(h_0 \gg h_a\) so that the constant density portion of the
shell will dominate the total column.  A small fraction \(e^{-\tau}\) of
the photons will penetrate the shell and be available to decelerate
the supersonic stream, as in the RBW case, yielding
\(v_0 = (1 - \tau e^{-\tau})\, v_\infty\).  In the illustrated case of
\(\tau = 3\), the resultant speed reduction before the shock is only 8\%.
As mentioned in \S~\ref{sec:effects-gravity}, radiative acceleration
becomes less important than gravity in the outer regions of an opaque
shell.  However, in order for this to have a significant effect on the
shell, the gravitational scale height would have to be be less than
the shell thickness: \(\Gamma\Edd h_a < h_0\).  From
equation~\eqref{eq:ha-versus-h0} this translates to
\(\tau > \frac43 \Gamma\Edd\), where Eddington factors are
\(\Gamma\Edd \sim 100\) for luminous stars.  Such large optical depths would
require extremely high ambient densities of
\(n > \num{2e7} L_4 \,v_{10}^2 \left( \SI{10}{M_\odot}/M \right)^2 \,
\si{cm^{-3}} \), where we have used equations~(\ref{eq:tau-star},
\ref{eq:eddington-factor}) and that \(\tau \approx 2\tau_*\) in the RBS limit.

@@

\subsection{Environments of bow-forming stars}
\label{sec:popul-bow-prod}

The stellar bows modeled in this paper can be seen as either due to
the motion of a star through the interstellar medium, or due to the
motion of the interstellar medium that flows past the star.  The
first case corresponds to runaway stars \citep{Blaauw:1961a}, which
have been ejected from a binary system or stellar cluster
\citep{Hoogerwerf:2001a}, while the second case can be due to
photoevaporation and champagne flows in \hii{} regions
\citep{Tenorio-Tagle:1979a, Shu:2002a, Henney:2005a}, or to general
turbulent flows in the Galaxy
\citep[e.g.,][]{Ballesteros-Paredes:1999a}.  In this section, we
consider the stream densities and velocities expected in each of these
scenarios, and compare them with our predictions for the type of bows
that should result.

Although some runaway stars have velocities exceeding
\SI{100}{km.s^{-1}}, most are moving considerably slower, with a
median peculiar velocity of about \SI{30}{km.s^{-1}}
\citep{Tetzlaff:2011a}.  The higher velocity runaways are likely
produced by dynamical interactions in the center of young clusters
\citep{Gualandris:2004a}, whereas the supernova-induced dissolution of
binary systems is predicted to primarily produce ``walkaways'' with
even slower velocities of order \SI{10}{km.s^{-1}}
\citep{Renzo:2018a}.  Environmental flows are also expected to be of
order \SI{10}{km.s^{-1}}, which is a characteristic velocity
dispersion for warm neutral gas in the inner Galaxy
\citep{Marasco:2017a} and also a typical expansion velocity for
\ion{H}{i} shells \citep{Ehlerova:2005a}.  Internal velocity
dispersions within \hii{} regions are subsonic
(\SIrange{5}{10}{km.s^{-1}}) for small regions, such as the Orion
Nebula, containing one or a few ionizing stars
\citep{Arthur:2016a}. This increases slowly for more luminous regions
as \(\sigma \propto L^{1/4}\) \citep{Bordalo:2011a}, so giant star forming
complexes such as Carina (100 times more luminous than Orion) show
velocity dispersions of order \SI{20}{km.s^{-1}}
\citep{Damiani:2016a}.  Higher velocities of
\SIrange{30}{40}{km.s^{-1}} are reached in divergent photoevaporation
flows \citep{Dyson:1968a}, but this is achieved at expense of a lower
density.  Given all the above, we expect there to be many more bows
with relative stream velocity \(v_\infty \sim \SI{20}{km.s^{-1}}\) than bows
with \(v_\infty \sim \SI{100}{km.s^{-1}}\).

Turning now to the ambient density, if bow shock stars were randomly
sampling the volume of the Galactic disk, then we would expect the
average density to be \(< \SI{1}{cm^{-3}}\).  The dominant gas phase
\citep{Ferriere:2001a} near the Galactic mid-plane is the Warm Neutral
Medium with a volume filling fraction of 45\% \citep[Fig.~11
of][]{Kalberla:2009a} and an average density of \SI{0.9}{cm^{-3}} at
the Solar circle \citep[\S~4 of][]{Kalberla:2008a}.  Significant
fractions of the volume (\(\approx 20\%\) each) are also occupied by the
Warm Ionized Medium and Hot Ionized Medium, which have even lower
densities (\(\approx \SI{0.3}{cm^{-3}}\) and \SI{e-3}{cm^{-3}},
respectively) and which increasingly dominate the volume for heights
\(z > \SI{500}{pc}\) above the plane.  The denser Cold Neutral Medium
(\(n \approx \text{\SIrange{10}{100}{cm^{-3}}}\)) and Molecular Clouds
(\(n > \SI{1000}{cm^{-3}}\)) occupy much smaller volumes
(\(\approx 10\%\) and \(< 1\%\), respectively).  However, observations of
stellar bows are clearly biased towards these higher densities.

In part, this bias is due to bows being easier to detect in denser
environments.  Depending on the emission mechanism, the shell
luminosity will be proportional to the column density
(\(\propto n R_0\)) or volume emission measure
(\(\propto n^2 R_0^3\)), which are both increasing functions of
\(n\).\footnote{%
  From \S~\ref{sec:three-bow-regimes}, the bow size falls relatively
  slowly as \(R_0 \propto n^{-1/2}\) in the WBS and RBS cases, and is
  independent of \(n\) in the RBW case.}  Another contribution to the
high-density bias is simply that none but the highest velocity
high-mass stars can move far during their lifetime from the molecular
clouds where they were born.  Even a runaway star with
\(v = \SI{30}{km.s^{-1}}\), will travel \(< \SI{120}{pc}\) during an
O-star lifetime of \(< \SI{4}{Myr}\), which is of the same order as
the sizes of Giant Molecular Clouds.  Many observed stellar bows are
found in high mass star clusters associated with large \hii{} regions
\citep{Povich:2008a, Sexton:2015b}.  The ionized gas density in
Carina, for example, has an average value of
\(\approx \SI{100}{cm^{-3}}\) \citep{Oberst:2011a, Damiani:2016a}, although
with peaks \(> \SI{3000}{cm^{-3}}\) in photoevaporation flows from
embedded molecular globules \citep{Smith:2004a}.  In more compact
\hii{} regions, such as the Orion Nebula, ionized densities up to
\SI{e4}{cm^{-3}} are found on scales of about \(\SI{0.1}{pc}\)
\citep{Weilbacher:2015a, ODell:2017b}.

In \S~\ref{sec:depend-stell-type} we found that the bow is
wind-supported when \(\tau_*^2 < \eta\wind\), where \(\tau_*\) is a fiducial
optical depth (eq.~[\ref{eq:taustar-typical}]) and \(\eta\wind\) is the
radiative momentum efficiency of the stellar wind
(eq.~[\ref{eq:wind-eta-typical}]).  This translates into a maximum
stream density for wind-supported bows of
\newcommand\crit{\ensuremath{_{\text{max}}}}
\newcommand\vink{\ensuremath{_{\text{\tiny Vink}}}}
\begin{equation}
  \label{eq:ncrit}
  n\crit \approx \left(\SI{30}{cm^{-3}}\right)
  \left(  \frac{v} {\SI{10}{km.s^{-1}}}\right)^2
  \left( \frac{\eta\wind}{\eta\vink} \right)
  \left( \frac{\SI{600}{cm^2.g^{-1}}}{\kappa} \right)
  \ ,
\end{equation}
which is largely insensitive to spectral type between early-B and
early-O stars.  In this equation, we have expressed the wind
efficiency in terms of the Table~\ref{tab:stars} values,
\(\eta\vink\), which are calculated from the \citet{Vink:1999a} recipe,
and the UV dust opacity in terms of the standard ISM value adopted in
\S~\ref{sec:depend-stell-type}.

Therefore, runaway stars moving through the diffuse ISM with
\(n \sim \SI{1}{cm^{-3}}\) and \(v \sim \SI{30}{km.s^{-1}}\) are safely in
the WBS regime, having \(n \approx \num{0.004}\, n\crit\) for the default
wind and dust parameters.  On the other hand, slow-moving stars in
\hii{} regions of moderate density (\(n \sim \SI{100}{cm^{-3}}\)) can
easily have \(n > n\crit\), meaning that their bows will be radiation
supported.

For cases where the stellar wind or dust properties differ from our
assumed values, then these conclusions may change.  For instance,
there are a growing number of stars, for which very low mass-loss
rates have been diagnosed from UV, optical or infrared line profiles
\citep{Martins:2005b, Marcolino:2009a, Najarro:2011a, Martins:2012a,
  Smith:2017b}.  Most of these stars are late-type O~dwarfs, for which
the \citet{Vink:2000a} prescription predicts a mass-loss rate of
\num{3e-8} to \SI{e-7}{M_\odot.yr^{-1}}, whereas the observed values are
in the range \num{e-10} to \SI{3e-9}{M_\odot.yr^{-1}} -- a shortfall of 10
to 1000!  This ``weak wind problem'' is a far larger discrepancy than
can be explained by clumping effects.  A potential (partial)
resolution is to suppose that internal shocks near the base of the
wind, together with thermal conduction, heat a large fraction of the
gas to coronal temperatures, where it is no longer detectable by means
of standard wind diagnostic lines \citep{Lucy:2012a}.  However, in
cases where x-ray diagnostics have been used to trace the mass loss
this hot component (\(\mu\)~Col, \citealp{Huenemoerder:2012a};
\(\zeta\)~Oph, \citealp{Cohen:2014a}), rates that are still an order of
magnitude below the Vink values are found.  We therefore show in

\subsection{Impact of the weak-wind phenomenon}
\label{sec:impact-weak-wind}

@@

% For main-sequence B~stars, wind column densities are too low to
% reliably measure the terminal velocity from near ultraviolet P~Cygni
% profiles \citep{Prinja:1989a}, and so the values are theory-dependent
% \citep{Krticka:2014a} and hence more uncertain.  A further
% complication is the existence of a subset of OB stars with anomalously
% weak winds \citep{Puls:2008a}, which in some cases is related to the
% presence of strong (\(\sim \SI{1}{kG}\)) magnetic fields
% \citep{Oskinova:2011b}.

@@

\subsection{Diagnosing bow type from shell emission profile}
\label{sec:diagnosing-bow-type}

Given that the predicted shell density profile varies between
wind-supported and radiation-supported bows
(\S~\ref{sec:axial-structure-bow} and Fig.~\ref{fig:axial-structure}),
it is worth considering whether it may be possible to use observed
emission profiles to distinguish between the two.
